---
title: "Data Visualization by Healy, Chapter 3"
output:
  pdf_document: default
  html_document: default
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(gapminder)
library(here)
```

# Figure 3.4

## Scatterplot of GDP per capita and Life Expectancy

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point()
```

## Plot with a smoothed line with confidence interval

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_smooth()
```

## Plot with a smoothed line and points

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point() + geom_smooth()
```

## Plot with a log scale on x-axis

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point() + scale_x_log10()
```

## Plot with a log scale on x-axis and change the x-axis labels

```{r}
p <- ggplot(data = gapminder, mapping = aes(x=gdpPercap, y = lifeExp))

p + geom_point() + scale_x_log10(labels = scales::comma) + geom_smooth(method = "lm")
```

## Changing the color of the points to purple

```{r}
p <- ggplot(data = gapminder, mapping = aes(x=gdpPercap, y = lifeExp))

p + geom_point(color = "purple") + scale_x_log10(labels = scales::comma) +
  geom_smooth(method = "loess")
```

## Change transparency and add labels, titles, and a caption (reasonably polished plot)

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(color = "orange", alpha=0.3) + 
  scale_x_log10(labels = scales::comma) +
  geom_smooth(method = "lm") + 
  labs(x = "GDP per Capita", y = "Life Expectancy in Years", 
       title = "Economic Growth and Life Expectancy",
       subtitle = "Data points in country-years",
       caption = "Source: Gapminder")
```

## Add a third variable to mapping and fill in standard error ribbon

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, 
                                            color = continent, fill = continent,)) 

p + geom_point() + 
  geom_smooth() + 
  scale_x_log10()
```

## Use only one smoothing line but keep the color aesthetic

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = continent)) + 
  geom_smooth() +
  scale_x_log10()
```

## Save plot with labels, title, and caption using ggsave()

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(color = "orange", alpha=0.3) + 
  scale_x_log10(labels = scales::comma) +
  geom_smooth(method = "lm") + 
  labs(x = "GDP per Capita", y = "Life Expectancy in Years", 
       title = "Economic Growth and Life Expectancy",
       subtitle = "Data points in country-years",
       caption = "Source: Gapminder")

ggsave(here("figures", "gdp_vs_lifeExp.pdf"))
```

# Chapter 3 Exercises

## 1) geom_smooth before geom_point

### The line is hidden behind the points (i.e. the line is plotted first since geom_smooth() is written first and the points are plotted on top of the line since geom_point() is written second).

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_smooth() + geom_point()
```

# Chapter 3 Exercises

## 2) Change GDP per Capita to population

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = pop, y = lifeExp))

p + geom_point()
```

# Chapter 3 Exercises

## 3) Alternative scale mappings

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point() + scale_x_sqrt()

p + geom_point() + scale_x_reverse()

p + geom_point() + scale_y_sqrt()

p + geom_point() + scale_y_reverse()
```

# Chapter 3 Exercises

## 4) Color to year instead of continent

### We see gradations of the color blue.

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = year))
```

# Chapter 3 Exercises

## 5) Instead of color = year, try color = factor(year)

### This provides a legend showing different colors for each of the years when data was collected.

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = factor(year)))
```





