---
title: "Data Visualization by Healy, Chapter 4"
output: pdf
date: "`r format(Sys.time(), '%B, %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r, message=FALSE}
library(ggplot2)
library(gapminder)
```

# Figure 4.1

## There is an error in the book. The text discusses the use of lifeExp as the outcome variable but the code (and figure 4.1) use gdpPercap. As a result, the figure produced from the code chunk below does not match Figure 4.1.

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp))

p + geom_line()
```

# Figure 4.1 follow up (subset data using only the United States and then create line graph with points).

```{r}
gapminder_data_us <- subset(gapminder, country=="United States")

p <- ggplot(data = gapminder_data_us, mapping = aes(x = year, y = lifeExp))

p + geom_line() + geom_point()
```

# Figure 4.2

## Here we use the group aesthetic to let ggplot know that the years are grouped by country.

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp))

p + geom_line(aes(group=country))
```

# Figure 4.3

## Faceting

### Faceting allows us to make "small multiple" plots separated out by an additional variable. This allows to convey a lot information to be presented compactly. A separate panel is drawn for each value of the faceting variable. Facets are not geoms but are a way of organizing geoms.

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp))

p + geom_line(aes(group = country)) + facet_wrap(~ continent)
```




































